---
title: "Trade-offs"
author: "John McMullen"
date: "2022-11-30"
output: html_document
editor_options: 
  chunk_output_type: console
---

Trade-offs between rhizobia over time and space

##Libraries requried
```{r}
##Libraries required

#ANOVA test associated packages
library(car) #Anova functions
library(emmeans) #post hoc testing
library(multcomp) #letter ranking
library(multcompView) #letter ranking
library(lme4) #mixed effect modeling
library(MuMIn) #AIC for quasi modeling

#beta regression
library(betareg) #beta regression test for proportional data
library(lmtest) #omnibus test

#Nonlinear regression
library(nlme) #gnls to fit logistic growth model by treatment
library(nlraa) #predict_gnls function to obtain estimate and confidence intervals for plotting

#PCA associated packages
library(vegan) #PCA
library(corrplot) #corrplot.mixed function
library(PerformanceAnalytics) #chart.Correlation
library(psych) #Bartlett's test and KMO

#data wrangling and plotting
library(dplyr) #summarize data
library(ggplot2) #plotting
```


##Data
```{r}
#data matrix (long format)
dat = read.csv(file = "~/Dropbox/docs/Lennon_Lab/Rhizobia_GEMS/Data/trade_offs/TO_pool.csv")

#convert to dummy variables
dat$X461 = as.factor(dat$X461)
dat$X262 = as.factor(dat$X262)

#convert to categorial data
dat$w_avg = as.factor(dat$w_avg)

#reorganize order for treatments
dat$trt = factor(dat$trt,levels= c("Apo","262","461","Mix"))
```


##Functions made
```{r}
#CI function for 95% confidence intervals
CI = function(x) {
  1.96*(sd(x)/length(x))
}

#CI function used to calculate growth rate CIs derived from a multivariate normal distribution
summarizer <- function(dat, alpha){
	n <- length(dat)
	quantiles <- c(alpha/2, 1-(alpha/2))
	CIs <- as.data.frame(matrix(NA, ncol(dat[[1]]), n*2))
	names(CIs) <- paste(rep(names(dat), each = 2), c("lo", "hi"), sep = ".")
	for(i in 1:n){
		CIs[,(2*i-1):(2*i)] <- t(apply(dat[[i]],    2, quantile, quantiles, na.rm = T))
		}
	return(CIs)
	}
```


##Plant traits
In our experiment, we scored several different plant traits, including leaf count (plant developmental age), plant biomass (developmental quality), stolon count (asexual reproduction), and chlorophyll content (nitrogen budget proxy).

###Biomass related analyses

####Above ground dry biomass
```{r}
#calculate the mean and CI 
biomass_raw = dat %>% group_by(w_avg,X461,X262,trt) %>% summarise(mean = mean(p_dry_mg),l.CI = mean(p_dry_mg)-CI(p_dry_mg),u.CI = mean(p_dry_mg)+CI(p_dry_mg))
biomass_raw$trt = factor(biomass_raw$trt,levels = c("Apo","262","461","Mix"))

#Plot the mean and CI over the categorical harvest weeks
ggplot(as.data.frame(biomass_raw), aes(x=trt,y=mean)) +
  geom_point(size=5,position = position_dodge(width=0.75))+
  geom_errorbar(aes(ymin=l.CI,ymax=u.CI),width=0.2,position = position_dodge(width=0.75)) +
  ylab(expression(atop("Dry biomass [mg]",paste("(mean \U00B1 CI, n = 5)")))) +
  facet_grid(.~w_avg) +
  theme_bw(base_size = 30) +
  theme(panel.grid = element_blank(),
        panel.border = element_rect(size = 1, colour = "black",fill=NA),
        panel.background = element_rect(fill="white"),
        axis.text.x=element_blank(),
        axis.text.y=element_text(size=rel(0.9)),
        axis.title.x = element_blank())
#############################

#Assess the effect of rhizobia presence over time points. Due to variation in magnitude of plant biomass, data was log10 transformed. The harvest date was included as a random effect initially.

biomass_date = lmer(log10(p_dry_mg) ~ X461*X262*w_avg + (1|w_est),data=dat)
biomass = lm(log10(p_dry_mg) ~ X461*X262*w_avg,data=dat)
anova(biomass_date,biomass) #no effect so dropped

summary(biomass)
AIC(biomass)

#assess normality of residuals
biomass_res = residuals(biomass)
summary(biomass_res) #mean and median are close
qqPlot(biomass_res) #falls within 95% CI
hist(biomass_res) #unimodal and fairly gaussian looking
boxplot(biomass_res) #fairly even spread of data
shapiro.test(biomass_res) #not significant
#All looks okay

#assess homoscedasticity of resdiuals
plot(fitted(biomass),biomass_res);
abline(h=0,lty=2);
lines(smooth.spline(fitted(biomass),biomass_res)); #fairly even spread
leveneTest(biomass_res ~ X461*X262*w_avg, data = dat) #not significant
#All looks okay

#omnibus test
Anova(biomass)

#post hoc by week
biomass_em = emmeans(biomass, ~ X262*X461|w_avg)
pairs(biomass_em)
as.data.frame(multcomp::cld(biomass_em))$.group

#############################

#setup model output to plot
biomass_plot = as.data.frame(biomass_em)
biomass_plot$trt = rep('A')
for(i in 1:nrow(biomass_plot)) {
  ifelse(biomass_plot[i,1]%in%"0" & biomass_plot[i,2]%in%"1",biomass_plot[i,9]<-"HQ",
  ifelse(biomass_plot[i,1]%in%"1" & biomass_plot[i,2]%in%"0",biomass_plot[i,9]<-"LQ",
  ifelse(biomass_plot[i,1]%in%"1" & biomass_plot[i,2]%in%"1",biomass_plot[i,9]<-"LQ:HQ",biomass_plot[i,9]<-"A")))
}
biomass_plot$trt = factor(biomass_plot$trt,levels = c("A","LQ","HQ","LQ:HQ"))

ggplot(biomass_plot, aes(x=trt,y=emmean)) +
  geom_point(size=5)+
  geom_errorbar(aes(ymin=emmean-SE,ymax=emmean+SE),width=0.3) +
  ylab(expression(atop(~Log[10]~"(biomass [mg])",paste("(mean \U00B1 SE, n = 5)")))) +
  scale_y_continuous(limits=c(0.85,4),breaks=seq(1,4,by=1))+
  facet_grid(.~w_avg) +
  theme_bw(base_size = 30) +
  theme(panel.grid = element_blank(),
        panel.border = element_rect(size = 1, colour = "black",fill=NA),
        panel.background = element_rect(fill="white"),
        axis.text.x=element_text(size=rel(0.9),angle=45,h=1),
        axis.text.y=element_text(size=rel(0.9)),
        axis.title.x = element_blank())
```

####Above ground fresh weight
```{r}
#calculate the mean and CI 
wet_raw = dat %>% group_by(w_avg,X461,X262,trt) %>% summarise(mean = mean(p_dry_mg),l.CI = mean(p_dry_mg)-CI(p_dry_mg),u.CI = mean(p_dry_mg)+CI(p_dry_mg))
wet_raw$trt = factor(wet_raw$trt,levels = c("Apo","262","461","Mix"))

#Plot the mean and CI over the categorical harvest weeks
ggplot(as.data.frame(wet_raw), aes(x=trt,y=mean)) +
  geom_point(size=5,position = position_dodge(width=0.75))+
  geom_errorbar(aes(ymin=l.CI,ymax=u.CI),width=0.2,position = position_dodge(width=0.75)) +
  ylab(expression(atop("Wet biomass [mg]",paste("(mean \U00B1 CI, n = 5)")))) +
  facet_grid(.~w_avg) +
  theme_bw(base_size = 30) +
  theme(panel.grid = element_blank(),
        panel.border = element_rect(size = 1, colour = "black",fill=NA),
        panel.background = element_rect(fill="white"),
        axis.text.x=element_blank(),
        axis.text.y=element_text(size=rel(0.9)),
        axis.title.x = element_blank())

#plot wet biomass by time (d)
ggplot(as.data.frame(dat), aes(x=d_post_inoc,y=p_wet,color=trt)) +
  geom_point(size=5)+
  geom_smooth(method='loess') +
  ylab(expression("Wet biomass [mg]")) +
  theme_bw(base_size = 30) +
  theme(panel.grid = element_blank(),
        panel.border = element_rect(size = 1, colour = "black",fill=NA),
        panel.background = element_rect(fill="white"),
        axis.text.x=element_text(size=rel(0.9)),
        axis.text.y=element_text(size=rel(0.9)),
        axis.title.x = element_blank())

#############################

#Assess the effect of rhizobia presence over time points. Due to variation in magnitude of plant biomass, data was log10 transformed. The harvest date was included as a random effect initially.

wet_date = lmer(log10(p_wet) ~ X461*X262*w_avg + (1|w_est),data=dat)
wet = lm(log10(p_wet) ~ X461*X262*w_avg,data=dat)
anova(wet_date,wet) #no effect so dropped

summary(wet)
AIC(wet)

#assess normality of residuals
wet_res = residuals(wet)
summary(wet_res) #mean and median are close
qqPlot(wet_res) #falls within 95% CI
hist(wet_res) #unimodal and fairly Gaussian looking
boxplot(wet_res) #fairly even spread of data
shapiro.test(wet_res) #not significant
#All looks okay

#assess homoscedasticity of residuals
plot(fitted(wet),wet_res);
abline(h=0,lty=2);
lines(smooth.spline(fitted(wet),wet_res)); #fairly even spread
leveneTest(wet_res ~ X461*X262*w_avg, data = dat) #not significant
#All looks okay

#omnibus test
Anova(wet)

#post hoc by week
wet_em = emmeans(wet, ~ X262*X461|w_avg)
pairs(wet_em)
multcomp::cld(wet_em)

#############################

#setup model output to plot
wet_plot = as.data.frame(wet_em)
wet_plot$trt = rep('A')
for(i in 1:nrow(wet_plot)) {
  ifelse(wet_plot[i,1]%in%"0" & wet_plot[i,2]%in%"1",wet_plot[i,9]<-"HQ",
  ifelse(wet_plot[i,1]%in%"1" & wet_plot[i,2]%in%"0",wet_plot[i,9]<-"LQ",
  ifelse(wet_plot[i,1]%in%"1" & wet_plot[i,2]%in%"1",wet_plot[i,9]<-"LQ:HQ","Apo")))
}

ggplot(wet_plot, aes(x=trt,y=emmean)) +
  geom_point(size=5)+
  geom_errorbar(aes(ymin=emmean-SE,ymax=emmean+SE),width=0.2) +
  ylab(expression(atop(Log[10]~"(fresh weight [mg])",paste("(mean \U00B1 SE, n = 5)")))) +
  scale_y_continuous(limits=c(1.7,5),breaks=seq(1,5,by=1))+
  facet_grid(.~w_avg) +
  theme_bw(base_size = 30) +
  theme(panel.grid = element_blank(),
        panel.border = element_rect(size = 1, colour = "black",fill=NA),
        panel.background = element_rect(fill="white"),
        axis.text.x=element_text(size=rel(0.9),angle=45,h=1),
        axis.text.y=element_text(size=rel(0.9)),
        axis.title.x = element_blank())
```

####Correlation between wet and dry biomass
```{r}
#plot dry by wet biomass to assess linearity
ggplot(data=dat,aes(x=(p_wet),y=(p_dry_mg))) +
  geom_point() +
  stat_smooth(method="lm")
#looks fairly linear in relationship

#Pearson's correlation
corr.test(x=dat$p_wet,y=dat$p_dry_mg,method="pearson") #highly correlated
corr.test(x=dat$p_wet,y=dat$p_dry_mg,method="pearson")$p #exact p value
```

####Effect of rhizobia on clover growth rate parameters
```{r}
#plot biomass by time relationship to see relationships in raw data with loess
ggplot(as.data.frame(dat), aes(x=(d_post_inoc),y=log10(p_dry_mg),color=trt)) +
  geom_point(size=5)+
  geom_smooth(method='loess') +
  #facet_grid(trt~.)+
  theme_bw(base_size = 30) +
  theme(panel.grid = element_blank(),
        panel.border = element_rect(size = 1, colour = "black",fill=NA),
        panel.background = element_rect(fill="white"),
        axis.text.x=element_text(size=rel(0.9)),
        axis.text.y=element_text(size=rel(0.9)),
        axis.title.x = element_blank())

#############################

#Assess the effect of rhizobia presence over time points. Due to variation in magnitude of plant biomass, data was log10 transformed.

#First, subset the dataset by treatment to get the initial start values for each treatment to model in the 3-parameter logistic growth model
fit_Apo <- getInitial(log10(p_dry_mg) ~ SSlogis(d_post_inoc, Asym, xmid, scal), data = dat[dat$trt=="Apo",])
print(fit_Apo)
fit_461 <- getInitial(log10(p_dry_mg) ~ SSlogis(d_post_inoc, Asym, xmid, scal), data = dat[dat$trt=="461",])
print(fit_461)
fit_262 <- getInitial(log10(p_dry_mg) ~ SSlogis(d_post_inoc, Asym, xmid, scal), data = dat[dat$trt=="262",])
print(fit_262)
fit_Mix <- getInitial(log10(p_dry_mg) ~ SSlogis(d_post_inoc, Asym, xmid, scal), data = dat[dat$trt=="Mix",])
print(fit_Mix)
fit_full <- getInitial(log10(p_dry_mg) ~ SSlogis(d_post_inoc, Asym, xmid, scal), data = dat)
print(fit_full)

#Full 3-parameter logistic growth model with each parameter term varying by treatment
gr_mdl = gnls(log10(p_dry_mg) ~ SSlogis(d_post_inoc,Asym,xmid,scal),data = dat,start=list(Asym=c(fit_Apo[1],fit_262[1],fit_461[1],fit_Mix[1]),xmid=c(fit_Apo[2],fit_262[2],fit_461[2],fit_Mix[2]),scal=c(fit_Apo[3],fit_262[3],fit_461[3],fit_Mix[3])),params=list(Asym~trt+0,scal~trt+0,xmid~trt+0))
summary(gr_mdl)

#Goodness of fit check for overall model
summary(gr_mdl) # RSE = 0.22976 = log10(mg) > does that mean the RSE is 1.7 mg?
cor(log10(dat$p_dry_mg),predict(gr_mdl)) #r = 0.9736383

#assess normality of residuals
rgr_res = residuals(gr_mdl)
summary(rgr_res) #mean and median are close
qqPlot(rgr_res) #falls within 95% CI
hist(rgr_res) #unimodal and fairly Gaussian looking
boxplot(rgr_res) #fairly even spread of data
shapiro.test(rgr_res) #not significant
#All looks okay

#assess homoscedasticity of residuals
plot(fitted(gr_mdl),rgr_res);
abline(h=0,lty=2);
lines(smooth.spline(fitted(gr_mdl),rgr_res)); #fairly even spread
#All looks okay

#############################

#Obtain predicted biomass and CI over time to plot
#need to redo the model fit since the response variable can't be transformed in the model function for predict
y=log10(p_dry_mg)
gr_mdl2 = gnls(y ~ SSlogis(d_post_inoc,Asym,xmid,scal),data = dat,start=list(Asym=c(fit_Apo[1],fit_262[1],fit_461[1],fit_Mix[1]),xmid=c(fit_Apo[2],fit_262[2],fit_461[2],fit_Mix[2]),scal=c(fit_Apo[3],fit_262[3],fit_461[3],fit_Mix[3])),params=list(Asym~trt+0,scal~trt+0,xmid~trt+0))
new = data.frame(d_post_inoc = rep(seq(min(dat$d_post_inoc),max(dat$d_post_inoc),by=1),4),
                 trt = rep(unique(dat$trt),each=length(seq(min(dat$d_post_inoc),max(dat$d_post_inoc),by=1))))
predframe <- predict_gnls(gr_mdl2,newdata = new,interval = "confidence")
new = cbind(new,predframe)

predframe <- with(dat,expand.grid(d_post_inoc=unique(d_post_inoc),trt=levels(trt)))
predframe$x <- predict(gr_mdl,newdata=predframe)

#plot data
ggplot() +
  geom_point(data=dat,aes(x=d_post_inoc,y=log10(p_dry_mg),color=trt),size=5)+
  geom_smooth(data=new,aes(x=d_post_inoc,y=p_dry_mg<-Estimate,color=trt)) +
  geom_ribbon(data=new,aes(x=d_post_inoc,ymin=Q2.5,ymax=Q97.5,color=trt),alpha=0.1,linetype=2) +
  scale_y_continuous(limits=c(0.5,4),breaks=seq(0.5,4,by=0.5))+
  scale_x_continuous(limits=c(15,115),breaks=seq(15,115,by=20))+
  theme_bw(base_size = 30) +
  xlab(expression("Time [dpi]")) +
  ylab(expression(Log[10]~"(biomass[mg])"))+
  theme(panel.grid = element_blank(),
        panel.border = element_rect(size = 1, colour = "black",fill=NA),
        panel.background = element_rect(fill="white"),
        axis.text.x=element_text(size=rel(0.9)),
        axis.text.y=element_text(size=rel(0.9)))

#############################

#Assess each model parameter to test whether it significanlty differs among treatments

#Drop the inverse umax term
umax_null = gnls(log10(p_dry_mg) ~ SSlogis(d_post_inoc,Asym,xmid,scal),data = dat,start=list(Asym=c(fit_Apo[1],fit_262[1],fit_461[1],fit_Mix[1]),xmid=c(fit_Apo[2],fit_262[2],fit_461[2],fit_Mix[2]),scal=fit_full[3]),params=list(Asym~trt+0,scal~1,xmid~trt+0))
summary(umax_null)

anova(umax_null,gr_mdl) # p = 0.2294

#Drop the mid point term when plants reach umax
xmid_null = gnls(log10(p_dry_mg) ~ SSlogis(d_post_inoc,Asym,xmid,scal),data = dat,start=list(Asym=c(fit_Apo[1],fit_262[1],fit_461[1],fit_Mix[1]),xmid=fit_full[2],scal=c(fit_Apo[3],fit_262[3],fit_461[3],fit_Mix[3])),params=list(Asym~trt+0,scal~trt+0,xmid~1))
summary(xmid_null)

anova(xmid_null,gr_mdl) # p = 0.0087

#Drop the asymptote term
Asym_null = gnls(log10(p_dry_mg) ~ SSlogis(d_post_inoc,Asym,xmid,scal),data = dat,start=list(Asym=fit_full[1],xmid=c(fit_Apo[2],fit_262[2],fit_461[2],fit_Mix[2]),scal=c(fit_Apo[3],fit_262[3],fit_461[3],fit_Mix[3])),params=list(Asym~1,scal~trt+0,xmid~trt+0))
summary(Asym_null)

anova(Asym_null,gr_mdl) # p < 0.0001

#Obtain CI for each parameter to show among treatment differences

gr_ci = as.data.frame(intervals(gr_mdl)$coef) #normally distributed confidence intervals for each parameter
gr_ci$coef = c(rep('K',4),rep('umax',4),rep('xmid',4))
gr_ci$trt = rep(c('A','LQ','HQ','LQ:HQ'))

#K
ggplot(gr_ci[gr_ci$coef=="K",],aes(x=trt,y=est.))+
  geom_point(size=5)+
  geom_errorbar(aes(ymin=lower,ymax=upper),width=0.2) +
  scale_y_continuous(limits=c(1.9,4),breaks=seq(2,4,by=1))+
  theme_bw(base_size = 30) +
  ylab(expression(Log[10]~"(biomass[mg])"))+
  theme(panel.grid = element_blank(),
        panel.border = element_rect(size = 1, colour = "black",fill=NA),
        panel.background = element_rect(fill="white"),
        axis.text.x=element_text(size=rel(0.9)),
        axis.text.y=element_text(size=rel(0.9)),
        axis.title.x = element_blank())

#xmid
ggplot(gr_ci[gr_ci$coef=="xmid",],aes(x=trt,y=est.))+
  geom_point(size=5)+
  geom_errorbar(aes(ymin=lower,ymax=upper),width=0.2) +
  scale_y_continuous(limits=c(8,40),breaks=seq(10,40,by=10))+
  theme_bw(base_size = 30) +
  ylab(expression("Mean development time [d]"))+
  theme(panel.grid = element_blank(),
        panel.border = element_rect(size = 1, colour = "black",fill=NA),
        panel.background = element_rect(fill="white"),
        axis.text.x=element_text(size=rel(0.9)),
        axis.text.y=element_text(size=rel(0.9)),
        axis.title.x = element_blank())

#umax
ggplot(gr_ci[gr_ci$coef=="umax",],aes(x=trt,y=1/est.))+
  geom_point(size=5)+
  geom_errorbar(aes(ymin=1/lower,ymax=1/upper),width=0.2) +
  scale_y_continuous(limits=c(0.035,0.14),breaks=seq(0.04,0.14,by=0.02))+
  theme_bw(base_size = 30) +
  ylab(expression("Maximum growth rate"))+
  theme(panel.grid = element_blank(),
        panel.border = element_rect(size = 1, colour = "black",fill=NA),
        panel.background = element_rect(fill="white"),
        axis.text.x=element_text(size=rel(0.9)),
        axis.text.y=element_text(size=rel(0.9)),
        axis.title.x = element_blank())

#############################

#Obtain biomass adjusted relative growth rate

#Convert model parameters to obtain maximum growth rate (umax), carrying capacity (K), and intercept (b)
parm = as.data.frame(matrix(nrow=4,ncol=7))
base::colnames(parm) <- c("trt","Asym","scal","xmid","K","umax","b")

#obtain SSlogis model parameters
mdl_coefs = coef(gr_mdl)
for(i in 1:4) {
  x = as.matrix(unique(dat$trt))
  parm[i,1] <- x[i,] #add treatment name
  y = mdl_coefs[grepl(x[i,],names(mdl_coefs))] #grab treatment specific coefficients
  parm[i,2:4] <- y #add 2:Asym, 3:scal, and 4:xmid
}

#Convert SSlogis model parameters for model: (b*K)/(b+(K-b)*exp(-r*t))
for(i in 1:4) { #i = rows
  parm[i,5] <- parm[i,2] #K; formula: K = Asym
  parm[i,6] <- 1/parm[i,3] #umax; formula: umax = 1/scal
  parm[i,7] <- parm[i,2]/(1+exp(parm[i,4]/parm[i,3])) #b; formula: b = Asym/(1+exp(xmid/scal))
}

#Calculate the RGR standardized to mass and then will compare the RGR at the average biomass among treatments to correct for variation in initial size

t = seq(min(dat$d_post_inoc),max(dat$d_post_inoc),0.25) #time sequence by quarter day along the measured times
RGRm = as.data.frame(matrix(nrow=length(t)*4,ncol=6)) #matrix to save RGR calculations
base::colnames(RGRm) <- c("trt","time","biomass","RGRm","lwr","upr")
RGRm$trt <- c(rep(parm[1,1],length(t)),rep(parm[2,1],length(t)),rep(parm[3,1],length(t)),rep(parm[4,1],length(t))) #add treatment IDs
RGRm$time <- rep(t,4) #add time sequence

#Set seed for distribution sampling in the for loop
set.seed(124)

for(i in 1:4) {
  name = as.matrix(unique(dat$trt))[i] #grab treatment ID
  x_parms = parm[parm$trt%in%name,] #grab the treatment coefficients
  K = x_parms[,5] #carrying capacity
  umax = x_parms[,6] #maximum growth rate
  b = x_parms[,7] #intercept
  
  #grab 10000 random replicates from a multivariate normal distribution using the model coefficients and their covariance matrix
  cov <- gr_mdl$varBeta #grab the covariance matrix for the coefficients
  x_cov <- cov[grepl(name,row.names(cov)),grepl(name,base::colnames(cov))] #grab treatment specific covariates
  row.names(x_cov) <- c("Asym","scal","xmid") #rename
  base::colnames(x_cov) <- c("Asym","scal","xmid") #rename
  x_coef <- x_parms[,2:4] #Asym, scal, and xmid
  x <- y <- data.frame(rmvnorm(n=10000, mean=as.matrix(x_coef), sigma=x_cov))
  base::colnames(x) <- base::colnames(y) <- c("Asym","scal","xmid")
		x$K  <- y$Asym #carrying capacity
		x$umax  <- 1/y$scal #maximum growth rate
		x$b <- y$Asym/(1 + exp(y$xmid/y$scal)) #intercept
		MCI <- RGRmCI <- matrix(NA, ncol = length(t), nrow = nrow(x))
		
		#calcualte the biomass and RGRm at each time point using the randomly sampled parameters to calculate the CIs for RGRm
			for(a in 1:nrow(x)){
			  x.a  <- x[a,]
		  	base::colnames(x.a) <- base::colnames(x)
		  	MCI[a,]     <- (x.a$b*x.a$K)/(x.a$b+(x.a$K-x.a$b)*exp(-x.a$umax*t)) #biomass calculation
		  	RGRmCI[a,] <- x.a$umax*MCI[a,]*(1 - MCI[a,]/x.a$K) #RGRm calculation
			}
		CIs <- summarizer(list(RGRmCI),0.05) #summarize the CI for the RGRm at 0.05 alpha threshold 
		print(name)
		temp=matrix(c(1,390,779,1168,389,778,1167,1556),ncol=2,nrow=4)
		strt = temp[i,1]
		fin = temp[i,2]
		RGRm[strt:fin,5:6] <- CIs #add the lower and upper 95% confidence intervals
		
		#below, calcualtes the predicted biomass and their respecitve RGRm
  for(j in strt:fin) {
    times = RGRm[j,2]
    m = (b*K)/(b+(K-b)*exp(-umax*times)) #biomass
    RGRm[j,3] <- m #save biomass
    RGRm[j,4] <- umax*m*(1-m/K) #calculate RGRm
  }
}

#plot the RGR over estimated biomass
ggplot(data=RGRm,aes(x=biomass,y=RGRm,color=trt)) +
  geom_smooth() +
  geom_ribbon(aes(ymin=lwr,ymax=upr),alpha=0.1,linetype=2) +
  scale_x_continuous(limits=c(0.9,4),breaks=seq(1,4,by=1))+
  scale_y_continuous(limits=c(0,0.1),breaks=seq(0,0.1,by=0.02))+
  theme_bw(base_size = 30) +
  ylab(expression("Relative growth rate")) +
  xlab(expression(Log[10]~"(biomass[mg])"))+
  theme(panel.grid = element_blank(),
        panel.border = element_rect(size = 1, colour = "black",fill=NA),
        panel.background = element_rect(fill="white"),
        axis.text.x=element_text(size=rel(0.9),h=1),
        axis.text.y=element_text(size=rel(0.9)))

#Find the RGR at the mean estimated biomass
mean(RGRm$biomass) #2.574553 which is great than the max biomass for the apos, so subset within apo range
min(RGRm[RGRm$trt=="Apo",3]) #1.07597
max(RGRm[RGRm$trt=="Apo",3]) #2.083248
mean(RGRm[RGRm$biomass<2.083248&RGRm$biomass>1.07597,3]) #1.759541
mean(RGRm[RGRm$trt!="Apo",3]) #2.814247
sum(RGRm[RGRm$biomass<2.083248&RGRm$biomass>1.07597&RGRm$trt!="Apo",3])/sum(RGRm[RGRm$trt!="Apo",3]) #10.4%

RGRextract = RGRm[RGRm$biomass>1.6&RGRm$biomass<2.9,] %>%
  mutate(rate = RGRm) %>%
  group_by(trt) %>%
  summarise(rgr = predict(loess(rate~biomass),c(1.759541,2.814247)),LCI = predict(loess(lwr~biomass),c(1.759541,2.814247)),UCI = predict(loess(upr~biomass),c(1.759541,2.814247))) %>%
  mutate(biomass = c(1.759541,2.814247))

RGRextract$trt2 = c(rep("LQ",2),rep("HQ",2),rep("A",2),rep("LQ:HQ",2))
RGRextract$trt2 = factor(RGRextract$trt2,levels=c("A","HQ","LQ","LQ:HQ"))
RGRextract$slope = rep(c("Among-treatment","Rhizobia-specific"),4)

ggplot(na.omit(RGRextract),aes(x=trt2,y=rgr)) +
  geom_point(size=5)+
  geom_errorbar(aes(ymin=LCI,ymax=UCI),width=0.2) +
  scale_y_continuous(limits=c(0,0.1),breaks=seq(0,0.1,by=0.02))+
  facet_grid(.~slope,scales="free") +
  theme_bw(base_size = 30) +
  ylab(expression("Relative growth rate")) +
  xlab(expression(Log[10]~"(biomass[mg])"))+
  theme(panel.grid = element_blank(),
        panel.border = element_rect(size = 1, colour = "black",fill=NA),
        panel.background = element_rect(fill="white"),
        axis.text.x=element_text(size=rel(0.9)),
        axis.text.y=element_text(size=rel(0.9)))
```

####dry matter content
```{r}
#calcualte the mean and CI 
dmc_raw = dat %>% group_by(w_avg,X461,X262,trt) %>% summarise(mean = mean(p_dry_mg/p_wet),l.CI = mean(p_dry_mg/p_wet)-CI(p_dry_mg/p_wet),u.CI = mean(p_dry_mg/p_wet)+CI(p_dry_mg/p_wet))
dmc_raw$trt = factor(dmc_raw$trt,levels = c("Apo","262","461","Mix"))

#Plot the mean and CI over the categorical harvest weeks
ggplot(as.data.frame(dmc_raw), aes(x=trt,y=mean)) +
  geom_point(size=5,position = position_dodge(width=0.75))+
  geom_errorbar(aes(ymin=l.CI,ymax=u.CI),width=0.2,position = position_dodge(width=0.75)) +
  ylab(expression(atop("DMC",paste("(mean \U00B1 CI, n = 5)")))) +
  facet_grid(.~w_avg) +
  theme_bw(base_size = 30) +
  theme(panel.grid = element_blank(),
        panel.border = element_rect(size = 1, colour = "black",fill=NA),
        panel.background = element_rect(fill="white"),
        axis.text.x=element_text(size=rel(0.9)),
        axis.text.y=element_text(size=rel(0.9)),
        axis.title.x = element_blank())

#plot dmc biomass by time (d)
ggplot(as.data.frame(dat), aes(x=d_post_inoc,y=p_dry_mg/p_wet,color=trt)) +
  geom_point(size=5)+
  geom_smooth(method='loess') +
  ylab(expression("DMC")) +
  theme_bw(base_size = 30) +
  theme(panel.grid = element_blank(),
        panel.border = element_rect(size = 1, colour = "black",fill=NA),
        panel.background = element_rect(fill="white"),
        axis.text.x=element_text(size=rel(0.9)),
        axis.text.y=element_text(size=rel(0.9)),
        axis.title.x = element_blank())

#############################

#Assess the effect of rhizobia presence over time points. 
x = dat$p_dry_mg/dat$p_wet
dmc = betareg(x ~ X461*X262*w_avg,weights = log10(p_wet),data=dat)
 
summary(dmc)
AIC(dmc)

#assess normality of residuals
dmc_res = residuals(dmc)
summary(dmc_res) #mean and median are close
qqPlot(dmc_res) #falls within 95% CI
hist(dmc_res) #unimodal and fairly gaussian looking
boxplot(dmc_res) #fairly even spread of data
#All looks okay

#assess homoscedasticity of resdiuals
plot(fitted(dmc),dmc_res);
abline(h=0,lty=2);
lines(smooth.spline(fitted(dmc),dmc_res)); #fairly even spread
#All looks okay

#omnibus test
lrtest(dmc)
joint_tests(dmc)

#post hoc by week
dmc_em = emmeans(dmc, ~ X262*X461|w_avg)
pairs(dmc_em)
multcomp::cld(dmc_em)

#setup data to plot
dmc_plot = as.data.frame(dmc_em)
dmc_plot$trt = rep('A')
for(i in 1:nrow(dmc_plot)) {
  ifelse(dmc_plot[i,1]%in%"0" & dmc_plot[i,2]%in%"1",dmc_plot[i,9]<-"HQ",
  ifelse(dmc_plot[i,1]%in%"1" & dmc_plot[i,2]%in%"0",dmc_plot[i,9]<-"LQ",
  ifelse(dmc_plot[i,1]%in%"1" & dmc_plot[i,2]%in%"1",dmc_plot[i,9]<-"LQ:HQ","A")))
}

ggplot(dmc_plot, aes(x=trt,y=emmean)) +
  geom_point(size=5)+
  geom_errorbar(aes(ymin=emmean-SE,ymax=emmean+SE),width=0.2) +
  ylab(expression(atop("Dry matter content",paste("(mean \U00B1 SE, n = 5)")))) +
  scale_y_continuous(limits=c(0.1,0.44),breaks=seq(0.1,0.4,by=0.1))+
  facet_grid(.~w_avg) +
  theme_bw(base_size = 30) +
  theme(panel.grid = element_blank(),
        panel.border = element_rect(size = 1, colour = "black",fill=NA),
        panel.background = element_rect(fill="white"),
        axis.text.x=element_text(size=rel(0.9),angle=45,h=1),
        axis.text.y=element_text(size=rel(0.9)),
        axis.title.x = element_blank())
```

###Leaf count related analyses

####Total leaf count
```{r}
#calcualte the geometric mean and geometric CI 
leaf_raw = dat %>% group_by(w_avg,X461,X262,trt) %>% summarise(mean = mean(log(leaf_T)),l.CI = mean(log(leaf_T))-CI(log(leaf_T)),u.CI = mean(log(leaf_T))+CI(log(leaf_T)))
leaf_raw$trt = factor(leaf_raw$trt,levels = c("Apo","262","461","Mix"))

#Plot the geometric mean and CI over the categorical harvest weeks
ggplot(as.data.frame(leaf_raw), aes(x=trt,y=exp(mean))) +
  geom_point(size=5,position = position_dodge(width=0.75))+
  geom_errorbar(aes(ymin=exp(l.CI),ymax=exp(u.CI)),width=0.2,position = position_dodge(width=0.75)) +
  ylab(expression(atop("Leaf count",paste("(mean \U00B1 CI, n = 5)")))) +
  facet_grid(.~w_avg) +
  theme_bw(base_size = 30) +
  theme(panel.grid = element_blank(),
        panel.border = element_rect(size = 1, colour = "black",fill=NA),
        panel.background = element_rect(fill="white"),
        axis.text.x=element_blank(),
        axis.text.y=element_text(size=rel(0.9)),
        axis.title.x = element_blank())

#plot leaf by time (d)
ggplot(as.data.frame(dat), aes(x=d_post_inoc,y=leaf_T,color=trt)) +
  geom_point(size=5)+
  geom_smooth(method='loess') +
  ylab(expression("Leaf count")) +
  theme_bw(base_size = 30) +
  theme(panel.grid = element_blank(),
        panel.border = element_rect(size = 1, colour = "black",fill=NA),
        panel.background = element_rect(fill="white"),
        axis.text.x=element_text(size=rel(0.9)),
        axis.text.y=element_text(size=rel(0.9)),
        axis.title.x = element_blank())

#Assess the effect of rhizobia presence over time points. 

hist(dat$leaf_T) #Poisson like distribution

#leaf = glm(leaf_T ~ X461*X262*w_avg,family=poisson(link="log"),data=dat)
leaf = glm(leaf_T ~ X461*X262*w_avg,family=quasipoisson(link="log"),data=dat) #Evidence of overdispersion

pchisq(summary(leaf)$dispersion * leaf$df.residual, leaf$df.residual, lower = F) # significance for overdispersion; p = 1.253203e-16

summary(leaf)

#omnibus test
Anova(leaf,test.statistic = "Wald")

#post hoc by week
leaf_em = emmeans(leaf, ~ X262*X461|w_avg,type="response")
pairs(leaf_em)
multcomp::cld(leaf_em)

#setup data to plot
leaf_plot = as.data.frame(leaf_em)
leaf_plot$trt = rep('A')
for(i in 1:nrow(leaf_plot)) {
  ifelse(leaf_plot[i,1]%in%"0" & leaf_plot[i,2]%in%"1",leaf_plot[i,9]<-"HQ",
  ifelse(leaf_plot[i,1]%in%"1" & leaf_plot[i,2]%in%"0",leaf_plot[i,9]<-"LQ",
  ifelse(leaf_plot[i,1]%in%"1" & leaf_plot[i,2]%in%"1",leaf_plot[i,9]<-"LQ:HQ",leaf_plot[i,9]<-"Apo")))
}
leaf_plot$trt = factor(leaf_plot$trt,levels = c("Apo","LQ","HQ","LQ:HQ"))


ggplot(leaf_plot, aes(x=trt,y=rate)) +
  geom_point(size=5)+
  geom_errorbar(aes(ymin=asymp.LCL,ymax=asymp.UCL),width=0.3) +
  ylab(expression(atop("Leaf count",paste("(mean \U00B1 CI, n = 5)")))) +
  #scale_y_continuous(limits=c(0.85,4),breaks=seq(1,4,by=1))+
  facet_grid(.~w_avg) +
  theme_bw(base_size = 30) +
  theme(panel.grid = element_blank(),
        panel.border = element_rect(size = 1, colour = "black",fill=NA),
        panel.background = element_rect(fill="white"),
        axis.text.x=element_text(size=rel(0.9),angle=45,h=1),
        axis.text.y=element_text(size=rel(0.9)),
        axis.title.x = element_blank())
```

#### Plant development time
```{r}
#plot biomass by time relationship between time and biomass
ggplot(as.data.frame(dat), aes(x=(d_post_inoc),y=(leaf_T),color=trt)) +
  geom_point(size=5)+
  geom_smooth(method='loess') +
  #facet_grid(trt~.)+
  theme_bw(base_size = 30) +
  theme(panel.grid = element_blank(),
        panel.border = element_rect(size = 1, colour = "black",fill=NA),
        panel.background = element_rect(fill="white"),
        axis.text.x=element_text(size=rel(0.9)),
        axis.text.y=element_text(size=rel(0.9)),
        axis.title.x = element_blank())

#Assess the effect of rhizobia presence over time points. Due to variation in magnitude of plant leaf count, data was log10 transformed. Compared power law and polynomial function. Polynomial function was a much better fit by AIC (power:69 vs. poly:-106)

rdr = lm(log10(leaf_T) ~ X461*X262*poly(d_post_inoc,2,raw=TRUE),data=dat) #compare degree 1 coefficients for instantaneous RGR
#rdr = lm(log(leaf_T) ~ X461*X262*log(d_post_inoc),data=dat) #power law to describe the change in RGR over time

summary(rdr)
AIC(rdr)

#assess individual fit for each treatment for polynomial regression
summary(lm(log10(leaf_T) ~ poly(d_post_inoc,2,raw=TRUE),data=dat[dat$trt=="Apo",])) #R2=0.9264
summary(lm(log10(leaf_T) ~ poly(d_post_inoc,2,raw=TRUE),data=dat[dat$trt=="461",])) #R2=0.9415
summary(lm(log10(leaf_T) ~ poly(d_post_inoc,2,raw=TRUE),data=dat[dat$trt=="262",])) #R2=0.8984
summary(lm(log10(leaf_T) ~ poly(d_post_inoc,2,raw=TRUE),data=dat[dat$trt=="Mix",])) #R2=0.9474

#assess normality of residuals
rdr_res = residuals(rdr)
summary(rdr_res) #mean and median are close
qqPlot(rdr_res) #falls within 95% CI
hist(rdr_res) #unimodal and fairly gaussian looking
boxplot(rdr_res) #fairly even spread of data
#All looks okay

#assess homoscedasticity of resdiuals
plot(fitted(rdr),rdr_res);
abline(h=0,lty=2);
lines(smooth.spline(fitted(rdr),rdr_res)); #fairly even spread
#All looks okay

#omnibus test
Anova(rdr,type="III")

#post hoc slope comparison to determine whether the presence of the other microbes influences the change in rdr
#summary(emtrends(rdr,specs=pairwise ~ X461|X262,var="log(d_post_inoc)",level = 0.975),infer=TRUE)
#summary(emtrends(rdr,specs=pairwise ~ X262|X461,var="log(d_post_inoc)",level = 0.975),infer=TRUE)
#summary(emtrends(rdr,specs=pairwise ~ X262+X461,var="log(d_post_inoc)"),infer=TRUE) #compare change in rdr
summary(emtrends(rdr,specs=pairwise ~ X262+X461,var="d_post_inoc",degree=1),infer=TRUE) #compare instantaneous rdr

#presence of the rhizobia induces similar changes in relative growth rate, while a lack of rhizobia results in slower relative growth rate.

#Plot fitted polynomial regression

ggplot(dat, aes(x=d_post_inoc,y=log10(leaf_T),color=trt)) +
  geom_point(size=5)+
  stat_smooth(method=lm,formula=y~poly(x,2,raw=TRUE)) +
  ylab(expression(log[10]~"(Leaf count)")) +
  scale_x_continuous(limits=c(0,120),breaks=seq(0,120,by=30))+
  scale_y_continuous(limits=c(0,3),breaks=seq(0,4,by=1))+
  theme_bw(base_size = 30) +
  theme(panel.grid = element_blank(),
        panel.border = element_rect(size = 1, colour = "black",fill=NA),
        panel.background = element_rect(fill="white"),
        axis.text.x=element_text(size=rel(0.9),angle=45,h=1),
        axis.text.y=element_text(size=rel(0.9)),
        axis.title.x = element_blank())

# Plot RGR (first degree marginal slopes) among treatments
RGR = as.data.frame(summary(emtrends(rgr,specs=pairwise ~ X262+X461,var="d_post_inoc",degree=1))$emtrends)
RGR$trt = rep('A')
for(i in 1:nrow(RGR)) {
  ifelse(RGR[i,1]%in%"0" & RGR[i,2]%in%"1",RGR[i,8]<-"HQ",
  ifelse(RGR[i,1]%in%"1" & RGR[i,2]%in%"0",RGR[i,8]<-"LQ",
  ifelse(RGR[i,1]%in%"1" & RGR[i,2]%in%"1",RGR[i,8]<-"LQ:HQ","Apo")))
}
RGR$trt = factor(RGR$trt,levels = c("A","HQ","LQ","LQ:HQ"))

ggplot(RGR, aes(x=trt,y=d_post_inoc.trend)) +
  geom_point(size=5)+
  geom_errorbar(aes(ymin=d_post_inoc.trend-SE,ymax=d_post_inoc.trend+SE),width=0.2) +
  ylab(expression("Relative growth rate")) +
  scale_y_continuous(limits=c(0.009,0.04),breaks=seq(0.01,.04,by=0.01))+
  theme_bw(base_size = 30) +
  theme(panel.grid = element_blank(),
        panel.border = element_rect(size = 1, colour = "black",fill=NA),
        panel.background = element_rect(fill="white"),
        axis.text.x=element_text(size=rel(0.9),angle=45,h=1),
        axis.text.y=element_text(size=rel(0.9)),
        axis.title.x = element_blank())
```

###Stolon count
```{r}
#calcualte the mean and CI 
stol_raw = dat %>% group_by(w_avg,X461,X262,trt) %>% summarise(mean = mean(stolon),l.CI = mean(stolon)-CI(stolon),u.CI = mean(stolon)+CI(stolon))
stol_raw$trt = factor(stol_raw$trt,levels = c("Apo","262","461","Mix"))

ggplot(as.data.frame(stol_raw), aes(x=trt,y=mean)) +
  geom_point(size=5,position = position_dodge(width=0.75))+
  geom_errorbar(aes(ymin=l.CI,ymax=u.CI),width=0.2,position = position_dodge(width=0.75)) +
  ylab(expression(atop("Stolon count",paste("(mean \U00B1 CI, n = 5)")))) +
  #xlab("Barcode presence") +
  #scale_y_continuous(limits=c(0.03,0.05),breaks=seq(0.03,0.05,by=0.01))+
  facet_grid(.~w_avg) +
  theme_bw(base_size = 30) +
  theme(panel.grid = element_blank(),
        panel.border = element_rect(size = 1, colour = "black",fill=NA),
        panel.background = element_rect(fill="white"),
        axis.text.x=element_blank(),
        axis.text.y=element_text(size=rel(0.9)),
        axis.title.x = element_blank())

# Compare the effect of rhizobia presence over each timepoint with poisson regression
hist(dat$stolon) #Poisson like distribution

#stol = glm(stolon+1 ~ X461*X262*w_avg,family=poisson(link="log"),data=dat)
stol = glm(stolon+1 ~ X461*X262*w_avg,family=quasipoisson(link="log"),data=dat) #evidence for underdispersion
summary(stol)
AIC(stol)

pchisq(summary(stol)$dispersion * stol$df.residual, stol$df.residual) # significance for dispersion; p = 2.077518e-05

#Wald's chi square test
Anova(stol,test.statistic = "Wald")

stol_em = emmeans(stol, ~ X262*X461|w_avg)
pairs(stol_em)
multcomp::cld(stol_em)

stol_plot = as.data.frame(stol_em)
stol_plot$trt = rep('A')
for(i in 1:nrow(stol_plot)) {
  ifelse(stol_plot[i,1]%in%"0" & stol_plot[i,2]%in%"1",stol_plot[i,9]<-"HQ",
  ifelse(stol_plot[i,1]%in%"1" & stol_plot[i,2]%in%"0",stol_plot[i,9]<-"LQ",
  ifelse(stol_plot[i,1]%in%"1" & stol_plot[i,2]%in%"1",stol_plot[i,9]<-"LQ:HQ","Apo")))
}
stol_plot$trt = factor(stol_plot$trt,levels = c("A","HQ","LQ","LQ:HQ"))

ggplot(stol_plot, aes(x=trt,y=emmean)) +
  geom_point(size=5)+
  geom_errorbar(aes(ymin=emmean-SE,ymax=emmean+SE),width=0.2) +
  ylab(expression(atop("dry plant stol",paste("(mean \U00B1 SE, n = 5)")))) +
  #scale_y_continuous(limits=c(0.03,0.05),breaks=seq(0.03,0.05,by=0.01))+
  facet_grid(.~w_avg) +
  theme_bw(base_size = 30) +
  theme(panel.grid = element_blank(),
        panel.border = element_rect(size = 1, colour = "black",fill=NA),
        panel.background = element_rect(fill="white"),
        axis.text.x=element_text(size=rel(0.9)),
        axis.text.y=element_text(size=rel(0.9)),
        axis.title.x = element_blank())
```

#Nodule mass
```{r}
#plot average and CI

nod_mass_raw = dat %>% group_by(w_avg,X461,X262,trt) %>% summarise(mean = mean(nod_avg_biomass_mg),l.CI = mean(nod_avg_biomass_mg)-CI(nod_avg_biomass_mg),u.CI = mean(nod_avg_biomass_mg)+CI(nod_avg_biomass_mg))
nod_mass_raw$trt = factor(nod_mass_raw$trt,levels = c("Apo","262","461","Mix"))

ggplot(subset(nod_mass_raw,w_avg!="16"&trt!="Apo"), aes(x=trt,y=mean)) +
  geom_point(size=5,position = position_dodge(width=0.75))+
  geom_errorbar(aes(ymin=l.CI,ymax=u.CI),width=0.2,position = position_dodge(width=0.75)) +
  ylab(expression(atop("Nodule biomass [mg "~nodule^-1~"]",paste("(mean \U00B1 CI, n = 5)")))) +
  #xlab("Barcode presence") +
  #scale_y_continuous(limits=c(0.03,0.05),breaks=seq(0.03,0.05,by=0.01))+
  facet_grid(.~w_avg) +
  theme_bw(base_size = 30) +
  theme(panel.grid = element_blank(),
        panel.border = element_rect(size = 1, colour = "black",fill=NA),
        panel.background = element_rect(fill="white"),
        axis.text.x=element_blank(),
        axis.text.y=element_text(size=rel(0.9)),
        axis.title.x = element_blank())


nod_mass = lm(log(nod_avg_biomass_mg) ~ trt*w_avg,weights=leaf_T,data=subset(dat,trt!="Apo"&w_avg!="16"))

summary(nod_mass)
AIC(nod_mass)

#assess normality
nod_mass_res = residuals(nod_mass)
summary(nod_mass_res)
qqPlot(nod_mass_res)
hist(nod_mass_res)
boxplot(nod_mass_res)
shapiro.test(nod_mass_res)

#assess homoscedasticity
plot(fitted(nod_mass),nod_mass_res);
abline(h=0,lty=2);
lines(smooth.spline(fitted(nod_mass),nod_mass_res));
leveneTest(nod_mass_res ~ X461*X262*w_avg, data = dat)
#looks okay

Anova(nod_mass,test.statistic = "F")

nod_mass_em = emmeans(nod_mass, ~ trt)
pairs(nod_mass_em)
multcomp::cld(nod_mass_em)

ggplot(as.data.frame(nod_mass_em), aes(x=X262+X461,y=emmean)) +
  geom_point(size=5)+
  geom_errorbar(aes(ymin=emmean-SE,ymax=emmean+SE),width=0.2) +
  ylab(expression(atop("dry plant nod_mass",paste("(mean \U00B1 SE, n = 5)")))) +
  scale_y_continuous(limits=c(0.03,0.05),breaks=seq(0.03,0.05,by=0.01))+
  facet_grid(.~w_avg) +
  theme_bw(base_size = 30) +
  theme(panel.grid = element_blank(),
        panel.border = element_rect(size = 1, colour = "black",fill=NA),
        panel.background = element_rect(fill="white"),
        axis.text.x=element_text(size=rel(0.9)),
        axis.text.y=element_text(size=rel(0.9)),
        axis.title.x = element_blank())
```

#Nodule CFU
```{r}
#plot average and CI

cfu_raw = dat %>% group_by(w_avg,X461,X262,trt) %>% summarise(mean = mean(CFU_nod),l.CI = mean(CFU_nod)-CI(CFU_nod),u.CI = mean(CFU_nod)+CI(CFU_nod))
cfu_raw$trt = factor(cfu_raw$trt,levels = c("Apo","262","461","Mix"))

ggplot(subset(cfu_raw,w_avg!="16"&trt!="Apo"), aes(x=trt,y=mean)) +
  geom_point(size=5,position = position_dodge(width=0.75))+
  geom_errorbar(aes(ymin=l.CI,ymax=u.CI),width=0.2,position = position_dodge(width=0.75)) +
  ylab(expression(atop("Nodule biomass [mg "~nodule^-1~"]",paste("(mean \U00B1 CI, n = 5)")))) +
  #xlab("Barcode presence") +
  #scale_y_continuous(limits=c(0.03,0.05),breaks=seq(0.03,0.05,by=0.01))+
  facet_grid(.~w_avg) +
  theme_bw(base_size = 30) +
  theme(panel.grid = element_blank(),
        panel.border = element_rect(size = 1, colour = "black",fill=NA),
        panel.background = element_rect(fill="white"),
        axis.text.x=element_blank(),
        axis.text.y=element_text(size=rel(0.9)),
        axis.title.x = element_blank())


cfu = lm(log10(CFU_nod) ~ trt*w_avg,weights=leaf_T,data=subset(dat,trt!="Apo"&w_avg!="16"))
cfu = lm(log(nod_avg_biomass_mg) ~ trt,weights=leaf_T,data=subset(dat,trt!="Apo"&w_avg!="16"))
summary(cfu)
AIC(cfu)

#assess normality
cfu_res = residuals(cfu)
summary(cfu_res)
qqPlot(cfu_res)
hist(cfu_res)
boxplot(cfu_res)
shapiro.test(cfu_res)

#assess homoscedasticity
plot(fitted(cfu),cfu_res);
abline(h=0,lty=2);
lines(smooth.spline(fitted(cfu),cfu_res));
leveneTest(cfu_res ~ X461*X262*w_avg, data = dat)
#looks okay

Anova(cfu,test.statistic = "F")

cfu_em = emmeans(cfu, ~ trt|w_avg)
pairs(cfu_em)
multcomp::cld(cfu_em)

ggplot(as.data.frame(cfu_em), aes(x=X262+X461,y=emmean)) +
  geom_point(size=5)+
  geom_errorbar(aes(ymin=emmean-SE,ymax=emmean+SE),width=0.2) +
  ylab(expression(atop("dry plant cfu",paste("(mean \U00B1 SE, n = 5)")))) +
  scale_y_continuous(limits=c(0.03,0.05),breaks=seq(0.03,0.05,by=0.01))+
  facet_grid(.~w_avg) +
  theme_bw(base_size = 30) +
  theme(panel.grid = element_blank(),
        panel.border = element_rect(size = 1, colour = "black",fill=NA),
        panel.background = element_rect(fill="white"),
        axis.text.x=element_text(size=rel(0.9)),
        axis.text.y=element_text(size=rel(0.9)),
        axis.title.x = element_blank())
```

#chlorophyll
```{r}
#plot average and CI

chlor_raw = dat %>% group_by(w_avg,X461,X262,trt) %>% summarise(mean = mean(chlor),l.CI = mean(chlor)-CI(chlor),u.CI = mean(chlor)+CI(chlor))
chlor_raw$trt = factor(chlor_raw$trt,levels = c("Apo","262","461","Mix"))

ggplot(as.data.frame(chlor_raw), aes(x=trt,y=mean)) +
  geom_point(size=5,position = position_dodge(width=0.75))+
  geom_errorbar(aes(ymin=l.CI,ymax=u.CI),width=0.2,position = position_dodge(width=0.75)) +
  ylab(expression(atop("Chlorophyll content",paste("(mean \U00B1 CI, n = 5)")))) +
  #xlab("Barcode presence") +
  #scale_y_continuous(limits=c(0.03,0.05),breaks=seq(0.03,0.05,by=0.01))+
  facet_grid(.~w_avg) +
  theme_bw(base_size = 30) +
  theme(panel.grid = element_blank(),
        panel.border = element_rect(size = 1, colour = "black",fill=NA),
        panel.background = element_rect(fill="white"),
        axis.text.x=element_blank(),
        axis.text.y=element_text(size=rel(0.9)),
        axis.title.x = element_blank())


chlor = lm(chlor ~ X461*X262*w_avg,data=dat)
summary(chlor)
AIC(chlor)

#assess normality
chlor_res = residuals(chlor)
summary(chlor_res)
qqPlot(chlor_res)
hist(chlor_res)
boxplot(chlor_res)
shapiro.test(chlor_res)

#assess homoscedasticity
plot(fitted(chlor),chlor_res);
abline(h=0,lty=2);
lines(smooth.spline(fitted(chlor),chlor_res));
leveneTest(chlor_res ~ X461*X262*w_avg, data = dat)
#looks okay

Anova(chlor,test.statistic = "F")

chlor_em = emmeans(chlor, ~ X262*X461|w_avg)
pairs(chlor_em)
multcomp::cld(chlor_em)

ggplot(as.data.frame(chlor_em), aes(x=X262+X461,y=emmean)) +
  geom_point(size=5)+
  geom_errorbar(aes(ymin=emmean-SE,ymax=emmean+SE),width=0.2) +
  ylab(expression(atop("dry plant chlor",paste("(mean \U00B1 SE, n = 5)")))) +
  scale_y_continuous(limits=c(0.03,0.05),breaks=seq(0.03,0.05,by=0.01))+
  facet_grid(.~w_avg) +
  theme_bw(base_size = 30) +
  theme(panel.grid = element_blank(),
        panel.border = element_rect(size = 1, colour = "black",fill=NA),
        panel.background = element_rect(fill="white"),
        axis.text.x=element_text(size=rel(0.9)),
        axis.text.y=element_text(size=rel(0.9)),
        axis.title.x = element_blank())
```

#photosynthetic potential
```{r}
#plot average and CI

photo_raw = dat %>% group_by(w_avg,X461,X262,trt) %>% summarise(mean = mean(prod),l.CI = mean(prod)-CI(prod),u.CI = mean(prod)+CI(prod))
photo_raw$trt = factor(photo_raw$trt,levels = c("Apo","262","461","Mix"))

ggplot(as.data.frame(photo_raw), aes(x=trt,y=mean)) +
  geom_point(size=5,position = position_dodge(width=0.75))+
  geom_errorbar(aes(ymin=l.CI,ymax=u.CI),width=0.2,position = position_dodge(width=0.75)) +
  ylab(expression(atop("photoophyll content",paste("(mean \U00B1 CI, n = 5)")))) +
  #xlab("Barcode presence") +
  #scale_y_continuous(limits=c(0.03,0.05),breaks=seq(0.03,0.05,by=0.01))+
  facet_grid(.~w_avg) +
  theme_bw(base_size = 30) +
  theme(panel.grid = element_blank(),
        panel.border = element_rect(size = 1, colour = "black",fill=NA),
        panel.background = element_rect(fill="white"),
        axis.text.x=element_blank(),
        axis.text.y=element_text(size=rel(0.9)),
        axis.title.x = element_blank())


photo = lm(log10(prod) ~ X461*X262*w_avg,data=dat)
summary(photo)
AIC(photo)

#assess normality
photo_res = residuals(photo)
summary(photo_res)
qqPlot(photo_res)
hist(photo_res)
boxplot(photo_res)
shapiro.test(photo_res)

#assess homoscedasticity
plot(fitted(photo),photo_res);
abline(h=0,lty=2);
lines(smooth.spline(fitted(photo),photo_res));
leveneTest(photo_res ~ X461*X262*w_avg, data = dat)
#looks okay

Anova(photo,test.statistic = "F")

photo_em = emmeans(photo, ~ X262*X461|w_avg)
pairs(photo_em)
multcomp::cld(photo_em)

ggplot(as.data.frame(photo_em), aes(x=X262+X461,y=emmean)) +
  geom_point(size=5)+
  geom_errorbar(aes(ymin=emmean-SE,ymax=emmean+SE),width=0.2) +
  ylab(expression(atop("dry plant photo",paste("(mean \U00B1 SE, n = 5)")))) +
  scale_y_continuous(limits=c(0.03,0.05),breaks=seq(0.03,0.05,by=0.01))+
  facet_grid(.~w_avg) +
  theme_bw(base_size = 30) +
  theme(panel.grid = element_blank(),
        panel.border = element_rect(size = 1, colour = "black",fill=NA),
        panel.background = element_rect(fill="white"),
        axis.text.x=element_text(size=rel(0.9)),
        axis.text.y=element_text(size=rel(0.9)),
        axis.title.x = element_blank())
```